Supplementary Material

### 2025-01-17 supplemental.tex

%% The amssymb package provides various useful mathematical symbols
% \usepackage{amssymb}
%% The amsmath package provides various useful equation environments.
% \usepackage{amsmath}
\usepackage{amsmath,amsthm,amssymb,scrextend,mathtools}
\usepackage{hyperref}
% \usepackage{newtxmath} %for making indicator function
% \RequirePackage[colorlinks,citecolor=blue,urlcolor=blue,backref=page,backref=page]{hyperref}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

@@

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}
\usepackage{natbib}
\usepackage{lmodern} %%helps deal with font size problems
\usepackage{dsfont}
\newcommand{\1}[1]{\mathds{1}\left[#1\right]}

@@

\begin{frontmatter}

@@

\title{Quantile Forecast Matching with a Bayesian Quantile Gaussian Process Model} %% Article title

@@

\author[label1]{Spencer Wadsworth} %% Author name
\author[label1]{Jarad Niemi}

%% Author affiliation
\affiliation[label1]{organization={Iowa State Univerity, Department of Statistics},%Department and Organization
            addressline={2438 Osborn Dr}, 
            city={Ames},
            postcode={50011}, 
            state={Iowa},
            country={USA}}

%% Abstract
\begin{abstract}
A set of probabilities along with corresponding quantiles are often used to 
define predictive distributions or probabilistic forecasts. These quantile 
predictions offer easily interpreted uncertainty of an event, 
and quantiles are generally straightforward to estimate using standard 
statistical and machine learning methods. However, compared to a distribution 
defined by a probability density or cumulative distribution function, a set of 
quantiles has less distributional information. When given estimated 
quantiles it may be desirable to estimate a fully defined continuous 
distribution function, and many researchers do so to make evaluation or 
ensemble modeling simpler. Most existing methods for fitting a distribution to 
quantiles lack accurate representation of the inherent uncertainty from 
quantile estimation or are limited in their applications. In this manuscript we 
present a Gaussian process model, the quantile Gaussian process, which is based 
on established theory of quantile functions and sample quantiles, to construct 
a probability distribution given estimated quantiles. A Bayesian application of 
the quantile Gaussian process is evaluated for parameter inference, and 
distribution approximation in simulation studies and the quantile Gaussian 
process is used to approximate the distributions of quantile forecasts from the 
2023-24 US Centers for Disease Control collaborative flu forecasting 
initiative. The simulation studies and data analysis show that the quantile 
Gaussian process leads to accurate inference on model parameters, estimation of 
a continuous distribution, and uncertainty quantification of sample quantiles.
\end{abstract}

@@

%%Research highlights
% \begin{highlights}
% \item Research highlight 1
% \item Research highlight 2
% \end{highlights}
% 
%% Keywords
\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
% 
% %% PACS codes here, in the form: \PACS code \sep code
% 
% %% MSC codes here, in the form: \MSC code \sep code
% %% or \MSC[2008] code \sep code (2000 is the default)
Sample quantiles \sep Quantile regression \sep Probabilistic forecasting \sep 
Disease outbreaks
\end{keyword}
% 
\end{frontmatter}

@@

\section{Introduction} \label{seq:intro}

The use of quantiles in statistical modeling and in reporting inferential or 
predictive uncertainty is widespread, and reporting several quantiles or 
predictive intervals is common for probabilistic forecasting 
\cite[]{gneiting2023model}. Quantiles can be easier to interpret than 
statistical model parameters and are often used to define confidence or 
prediction intervals. Thus with multiple quantiles for a particular outcome, 
one has a measure of uncertainty.
Estimating quantiles in the presence of covariates via 
quantile regression is a common
statistical and machine learning strategy, and where parametric models are 
complicated or nonexistent quantiles may be easier to estimate via machine 
learning methods \cite[]{martin2022direct,chung2021beyond, koenker2017quantile, 
koenker1978regression}. In estimating multiple quantiles, quantile regression 
often provides a tradeoff with parametric modeling in that for some problems 
quantile regression may be easier to perform but that predicted quantiles lack 
the detailed information of a fully defined predictive distribution 
\cite[]{pohle2020murphy}. 
% Quantile regression may also lead to other issues 
% such as quantile crossing where an estimated quantile may be smaller than 
% another estimated quantile with for a lower probability level 
% \cite[]{chernozhukov2010quantile,he1997quantile}.
% Data is often reported as quantiles, 
Perhaps, for data privacy reasons, data quantiles are often reported as 
summaries of the data.
% along with predictive distributions. 
% \jarad{Predictive distributions for what?}
Census or medical data, which can be very large or personal to the subjects, 
may be published as summary or aggregate data including percentiles (quantiles) 
and medians 
\cite[]{simpson2023interpolating,cdc2022growthcharts,nirwan2020bayesian}. In 
collaborative forecast initiatives and competitions, probabilistic forecasts 
are often submitted as a set of predictive quantiles 
% for given probabilities 
\cite[]{gneiting2023model,hong2016probabilistic}. In recent years disease 
outbreak forecast hubs require that all forecasts submitted by outside 
participants be represented by  several predictive intervals or quantiles. 
This standardized representation allows for straightforward forecast scoring 
and ensemble building 
\cite[]{mathis2024evaluation,mathis2023flusight,Cramer2022-hub-dataset,
cramer2022evaluation,sherratt2023predictive,bracher2021evaluating}.

A set of quantiles may provide distributional information for an event, but it 
is not as informative as a distribution defined by a cumulative distribution 
function (CDF), the inverse-CDF or quantile function QF, or a probability 
density/mass function (PDF), and quantiles provide no distribution tail 
information --information for values below the smallest quantile or above the 
largest quantile. Another drawback for using quantiles to define a distribution 
is that many tools for evaluating or scoring continuous distributions require 
a CDF or PDF
% \jarad{what does 'fully defined' mean?} 
\cite[]{gneiting2007strictly,gneiting2014probabilistic}. Likewise combining 
distributions into an ensemble distribution is commonly 
% \jarad{most often? perhaps use commonly}
done by aggregating multiple QFs or 
CDFs/PDFs, the latter being possible only when a CDF/PDF is available 
\cite[]{gneiting2005calibrated,wang2023forecast}.

Fitting quantiles to recover or estimate a continuous distribution is done in 
many fields, often for the purpose of evaluating forecasts using rules that 
require a CDF or PDF \cite[]{simpson2023interpolating, gerding2023evaluating}. 
Fitting may also be done to allow combining multiple forecasts using aggregation 
methods that require CDFs or PDFs 
\cite[]{gyamerah2020probabilistic,li2019combining,baran2018combining,
bogner2017combining,he2016short,gneiting2005calibrated}. 
An example of fitting quantiles is given in figure 
\ref{fig:quant_match_example}. The 12 points in the figure are quantiles 
estimated from a random sample of size 100 from a standard normal distribution 
for given probabilities. The quantile function of the standard normal 
distribution is represented by the solid grey line in the plot. 
The dashed grey line is the fitted QF of a normal 
distribution that was estimated by selecting the mean and standard deviation 
parameters which minimize the least squares distance between the estimated 
quantiles and the QF.

\begin{figure}[hbt!]
    \centering
    \includegraphics[scale=.5]{Images/fit_quantiles_example.png}
    \caption{Points representing 12 quantiles ($y$-axis) for given probability 
    values ($x$-axis) estimated from a random sample of size 100 from a 
    standard normal distribution with the true quantile function (QF) 
    (solid grey). The estimated QF (dashed grey) was fit by selecting the mean 
    and standard deviation 
    parameters which minimize the least squares distance between the quantiles 
    and the QF.}
    \label{fig:quant_match_example}
\end{figure}

Hereafter we refer to estimating a continuous distribution by fitting 
quantiles as quantile matching (QM).
\cite{sgouropoulos2015matching} performed QM by minimizing the mean square 
difference between quantiles of a response variable and a linear combination 
of quantiles of covariates. Selecting distribution parameters which minimize 
the mean square error
% \jarad{is least squares different than minimizing the mean square difference?}
between quantiles and a QF is a common QM method 
\cite[]{dilger2022distributions,li2019combining, belgorodski2017quantilemse}, 
% \jarad{Describe, briefly, what these methods do rather than just listing them.}
kernel density estimation and spline interpolation are common nonparametric 
QM methods 
\cite[]{gerding2023evaluating,gyamerah2020probabilistic,he2016short}, 
and \cite{keelin2016metalog} introduced a semiparametric method based on 
defining a flexible QF to fit quantiles. An issue shared by these methods is 
that they do not adequately account for the uncertainty inherent from the 
estimation of quantiles.  \cite{nirwan2020bayesian} introduced a QM model 
based on the definition of sample quantiles as order statistics and formulated
% \jarad{using as?}
a model likelihood for the quantiles as the joint PDF of multiple order 
statistics. This model 
provides exact inference for quantile uncertainty, but relies on the CDF and 
PDF of the distribution, making it less practical for fitting quantile defined 
distributions which have a QF but for which the CDF and PDF either do not exist 
in closed form or are difficult to evaluate 
\cite[]{perepolkin2023tenets,joiner1971some,tukey1960practical}.

In this manuscript we introduce a novel model, the quantile Gaussian process 
(QGP), used specifically for QM. The context for using the QGP is that the 
available data is a set of quantiles estimated for a given set of 
probabilities.
The QGP relies on established theory underlying the relationship between the 
QF of a continuous distribution and sample quantiles 
\cite[]{parzen2004quantile, gilchrist2000statistical,hyndman1996sample,
walker1968note,cramer1951mathematical}. 
The QGP can provide accurate asymptotic inference of the quantile uncertainty, 
and it works well for QM when modeling quantile defined distributions. 
In section \ref{sec:quant_func} the QF of a continuous random variable and 
sample quantiles are each defined, and important properties and asymptotic 
theory are reviewed. In section \ref{sec:qgp_model} the QGP model for QM is 
introduced. Section \ref{sec:simulation_analyses} contains simulation studies 
illustrating the QGP's ability to make parameter inference and match the 
distribution from which the quantiles where estimated. Section 
\ref{sec:qm_analysis} contains an application of QM using the QGP model for 
quantile forecasts of the 2023-24 United States Centers for Disease Control and
Prevention
(CDC) flu forecasting competition, also known as FluSight. Section 
\ref{sec:qm_conclusion} concludes this manuscript with a short summary and 
ideas for further development in QM modeling.

\section{Quantile function and sample quantiles} \label{sec:quant_func}

Along with the CDF and PDF, the QF is a defining function of a random variable. 
The QF, however, often receives less attention in standard statistical 
training than do the CDF and PDF \cite[]{parzen2004quantile}, and the CDF and 
PDF are more used often for modeling and inference.
% But the QF tends to be the little brother 
% \jarad{little brother?}
% of both the CDF and PDF in that the latter two functions are taught and used far more in statistical modeling and inference. 
% Nevertheless, the QF and several of its properties are central to the QGP model introduced in this paper. 
% \jarad{another throwaway sentence}
This section contains the definition of the QF, important properties, the 
definition of sample quantiles, and key central limit theorems (CLTs) of sample 
quantiles.

\subsection{Quantile function definition and properties}

The QF is defined in definition \ref{def:qf}.

\begin{definition}
    \label{def:qf}
    For a cumulative distribution function $F: \mathbb{R} \rightarrow 
    \left[ 0,1 \right]$, 
    the quantile function $Q: \left[0,1\right] 
    \rightarrow \mathbb{R}$ is defined as

    \begin{equation*}
    \label{eq:qf}
        Q\left(p\right) = F^{-1}\left(p\right) := 
        \emph{inf}\{x \in \mathbb{R} : F\left(x\right) \geq p\}, 
        \quad p \in \left[0,1\right]
    \end{equation*}
    An alternative definition not in terms of the cumulative distribution 
    function is a function 
    $Q: \left[0,1\right] \rightarrow \mathbb{R} \cup \{-\infty, \infty\}$, which is 
    nondecreasing and left-continuous. 
\end{definition}
% \jarad{the latter does not sound like a definition to me, but instead lists two properties.}

For most statistical applications, it is useful to define the QF as the inverse 
of the CDF. However, it is sometimes the case that defining a distribution by 
a QF rather than a CDF is advantageous. For instance when a distribution is 
defined by a 
QF but has no closed form CDF or when simple QFs may be aggregated to define 
more complex distributions 
\cite[]{perepolkin2023tenets, gasthaus2019probabilistic, alvarez2023quantile}. 
This section continues with important properties of the QF and functions 
including the location-scale property, the quantile density function, the 
probability integral transform, and examples of quantile defined distributions.
%(should the last reference be included??).

\subsubsection{Location-scale property}
For a random variable with PDF $f_0\left(x\right)$ and 
QF $Q_0\left(p\right)$ for $p \in \left[0,1\right]$, a 
location-scale family takes the PDF form 
$\left(1/\sigma\right)f_0\left(\left(x - \mu\right)/\sigma\right)$ 
with location parameter 
$\mu \in \left( -\infty, \infty \right)$ and scale parameter 
$\sigma \in \left(0, \infty \right)$ \cite[]{casella2002statistical}. 
The QF of the 
location-scale family then takes the form $\mu + \sigma Q_0\left(p \right)$ 
\cite[]{parzen2004quantile}. By the addition and multiplication rules from 
section 3.2 of \cite{gilchrist2000statistical}, 
%(should I include a table listing properties somehwere??), 
this is a valid QF. \cite{gilchrist2000statistical} outlines the 
standardization rule which states that if the random variable $Y$ with QF 
$Q_Y\left(p \right)$ has a standard distribution, that is the 
random variable $Y$ is 
centered (by a mean or median) at 0 and has a scale of 1, then a random 
variable with quantile function 
$Q_{T\left(Y \right)}\left(p \right) = \mu + \sigma Q_Y\left(p \right)$ is 
centered at $\mu$ and has scale $\sigma$.

The linear form for a QF is already inherent in some distributions families 
such as the normal and logistic distribution families, with respective QFs 
$\mu + \sigma \Phi^{-1} \left(p \right)$ and 
$\mu + \sigma \text{log}\left(p/ \left(1 - p \right) \right)$. For the
normal distribution, 
$\Phi^{-1}\left(p \right)$ is the QF of a standard normal distribution 
and $\mu$ and $\sigma$ are the mean and standard deviation of the 
location-scale transformed normal distribution. For the logistic distribution 
$\text{log}\left(p/\left(1-p \right) \right)$ 
is the QF of a logistic distribution with mean 0 and 
scale 1, and $\mu$ and $\sigma$ become the mean and scale parameters of the 
location-scale transformed logistic distribution. The linear form is convenient 
for quantile regression modeling and is key to the development of the metalog 
distribution family which was designed specifically for modeling quantiles 
\cite[]{keelin2016metalog}. 
% Two examples of quantile defined distributions  are give further in this section.

\subsubsection{Quantile density function}

Besides the CDF, PDF, and QF a  fourth infrequently mentioned defining function 
of a continuous random variable is the quantile density function (QDF). As the 
PDF is the derivative of the CDF, the QDF $q:[0,1] \rightarrow \mathbb{R}$ is 
the derivative of the QF defined as
\[
    q(p) = \frac{dQ(p)}{dp}.
\]
If $Q(p)$ is the inverse of a CDF $F(x)$, a calculus result shows that the 
reciprocal QDF $[q(p)]^{-1} = f(Q(p))$, or the PDF evaluated at the QF 
evaluated at $p$ 
\cite[]{perepolkin2023tenets, gilchrist2000statistical}. The QDF is important 
for establishing the CLTs in section \ref{sec:samp_quants} and for the QGP 
models introduced in section \ref{sec:qgp_model}.

\subsubsection{Probability integral transform}
An important application of the QF is using it to sample from continuous 
probability distributions via the probability integral transform (PIT). The 
PIT states that for any continuous random variable $Y$ with CDF $F_Y(y)$, the 
transformed random variable $X = F_Y(Y)$ is uniformly distributed on $(0,1)$, 
or $X \sim \text{Unif}(0,1)$. Thus if one can sample from a standard uniform 
distribution, one may also sample from a continuous distribution provided the 
QF of that distribution can be evaluated or reasonably approximated 
\cite[]{wilkinson2018stochastic}. The PIT is useful for assessing model fit via 
generalized residuals \cite[]{yang2024double, cox1968general} and the 
calibrating of probabilistic forecasts \cite[]{gneiting2007probabilistic}. 
Herein the PIT is important for developing one version of the QGP model, and it 
is used as part of a distance measure for assessing QM.

\subsubsection{Examples of quantile defined distributions} \label{sec:quant_def_dist}

An example of a quantile defined distribution family that lacks a CDF in closed 
form is the generalized lambda distribution (GLD) family 
\cite[]{ramberg1974approximate, perepolkin2023tenets}. A special case of the 
GLD family is the Tukey lambda distribution (TLD) family, a one parameter 
distribution family introduced by \cite{tukey1960practical}, 
\cite[]{joiner1971some}. The QF for the TLD is in (\ref{eq:tuk_lam}) and the 
QDF in (\ref{eq:tuk_lam_qd}).

\begin{equation}
    \label{eq:tuk_lam}
    Q_{\lambda}(p) = \begin{cases} 
      \frac{1}{\lambda} \left[p^{\lambda} - (1 - p)^{\lambda} \right], & \lambda \neq 0 \\
      \text{log} \left(\frac{p}{1 - p} \right), & \lambda = 0
   \end{cases}
\end{equation}

\begin{equation}
    \label{eq:tuk_lam_qd}
    q_{\lambda}(p) = p^{\lambda - 1} + (1-p)^{\lambda -1} 
\end{equation}

Another distribution family defined by its quantile function but without a 
closed form CDF is the metalog distribution (MLD) family, a generalization of 
the logistic distribution family \cite[]{keelin2016metalog}. One version of the 
MLD is defined in (\ref{eq:mldq}).

\begin{equation}
    \label{eq:mldq}
    Q_{\textbf{a}}(p) = a_1 + a_3(p - 0.5) + a_2\text{log}\left(\frac{p}{1-p} \right)
\end{equation}
These two distributions are used in section \ref{sec:simulation_analyses} 
for comparing different quantile matching methods.

\subsection{Sample quantiles} \label{sec:samp_quants}

Using samples of a random variable to estimate quantiles is prevalent in 
statistical modeling. Given a random sample from some distribution, the sample 
quantile function may be defined in terms of the order statistics. 
\cite{hyndman1996sample} review several definitions of sample quantiles used in 
statistical packages and state definition \ref{def:sqf} as a general definition 
of the functions reviewed.

\begin{definition}
For $n$ independent observations $Y_1, ..., Y_n$ from a distribution with 
corresponding order statistics $Y_{(1)}, ..., Y_{(n)}$, the sample quantile 
function $\hat{Q}_n(p)$ may be defined as
    \label{def:sqf}
    \begin{equation}
        \label{eq:quant_df}
        \hat{Q}_n(p) = (1- \gamma)Y_{(j)} + \gamma Y_{(j + 1)}
    \end{equation}
    where 
    \[\frac{j - m}{n} \leq p < \frac{j - m + 1}{n}
    \]
    for some $m \in \mathbb{R}$ and $0 \leq \gamma \leq 1$ where 
    $\gamma$ is some function of $j = \lfloor pn + m \rfloor$,  
    $g = pn + m - j$, and $\lfloor \cdot \rfloor$ is the floor function.
\end{definition}

Sample quantiles are used to give a probability summary of a dataset. Sample 
quantiles of Markov chain Monte Carlo draws from a Bayesian posterior 
distribution are used to analyze posterior distributions, for example by using 
quantiles to form credible intervals. 
Well known asymptotic results of sample quantiles are presented below, and it 
may be noted that asymptotic results for quantile regression quantiles are 
similar \cite[]{kocherginsky2005practical, koenker1978regression}. 
% (maybe rethink this statement or the specific citations??).

\subsubsection{Sample quantile central limit theorem}

The asymptotic distribution of a set of sample quantiles is given in 
(\ref{eq:qclt}), and it is the basis for the QGP model for QM in section 
\ref{sec:qgp_model}. For independent draws $Y_1, ..., Y_n$ from a continuous 
random variable with CDF $F$, PDF $f$, and QF $Q$, and with a set of sample 
quantiles calculated by (\ref{eq:quant_df}), theorem \ref{thm:qclt} holds. 
Note that in (\ref{eq:qclt}), $N_K(\cdot, \cdot)$ refers to a multivariate 
normal distribution with $K$ dimensions, and in (\ref{eq:qclt_covariance}) the 
operator $a \wedge b$ is the minimum of the values $a$ and $b$.

\begin{theorem}{\emph{Sample quantile central limit theorem}}
\label{thm:qclt}

Given a vector of length $K$ of probabilities 
$\boldsymbol{p} = (p_1, ..., p_K)$ and the corresponding quantile vector 
$\hat{\boldsymbol{Q}}_n(\boldsymbol{p}) = (\hat{Q}_n(p_1), ..., \hat{Q}_n(p_K))$, 
if $F$ is absolutely continuous for all $y \in \mathcal{Y}$ and is strictly 
increasing, then

\begin{equation}
    \label{eq:qclt}
    \sqrt{n}(\hat{\boldsymbol{Q}}_n(\boldsymbol{p}) - 
    \boldsymbol{Q}(\boldsymbol{p})) \overset{D}{\rightarrow} 
    N_K(0, \boldsymbol{\mathcal{K}})
\end{equation}
where $\boldsymbol{\mathcal{K}}$ is a $K\times K$ with the $ij^{th}$ 
entry $\kappa_{ij}$ where

\begin{equation}
    \label{eq:qclt_covariance}
    \kappa_{ij} = \frac{p_i \wedge p_j - p_i p_j}{f(Q(p_i)) f(Q(p_j))}
\end{equation}
\end{theorem}

Theorem \ref{thm:qclt} has been known since the mid 20th century 
\cite[]{cramer1951mathematical}, and a thorough though not unique proof is 
found in \cite{walker1968note}. The entries in the covariance matrix in 
(\ref{eq:qclt_covariance}) may equivalently be written as
\begin{equation}
    \rho_{ij} = [p_i \wedge p_j - p_i p_j]q(p_i)q(p_j)
    \label{eq:qgp_simp_cov}
\end{equation}
where $q$ is the QDF. Another CLT in (\ref{eq:tqclt}) for a set of quantiles 
transformed by the underlying CDF $F(\hat{\boldsymbol{Q}}_n(\boldsymbol{p})) = 
(F(\hat{Q}_n(p_1)), ..., F(\hat{Q}_n(p_k)))$ is in (\ref{eq:qclt}) and can be 
derived from (\ref{eq:qclt}) by using the PIT and the Delta method 
\cite[]{parzen2004quantile}.

\begin{equation}
    \label{eq:tqclt}
    \sqrt{n}(F(\hat{\boldsymbol{Q}}_n(\boldsymbol{p})) - \boldsymbol{p}) 
    \overset{D}{\rightarrow} N_K(0, \boldsymbol{\Gamma})
\end{equation}
Here the covariance matrix $\Gamma$ has entries $\Gamma_{ij} = p_i \wedge p_j - 
p_i p_j$, making the asymptotic distribution in (\ref{eq:tqclt}) a 
Brownian bridge \cite[]{chow2009brownian}. This second result makes QM using 
the QGP model possible where $Q(p)$ is difficult to solve as is shown in the 
following section.

\section{Quantile Gaussian process model} \label{sec:qgp_model}

We consider the situation where one is given a set of data including a vector 
of $K$ probabilities $\boldsymbol{p} = (p_1, ..., p_K)$ and a vector of 
estimated quantiles at the given probability levels 
$\hat{\boldsymbol{Q}}_n(\boldsymbol{p}) = (\hat{Q}_n(p_1), ..., \hat{Q}_n(p_K))$. 
The set of quantiles may provide useful information about a distribution, but 
the information is limited and one may desire a more complete distribution.
The QGP model for QM and estimating a continuous distribution based on the CLT 
in (\ref{eq:qclt}) is in (\ref{eq:qgp}).

\begin{equation}
    \label{eq:qgp}
\hat{\boldsymbol{Q}}_n(\boldsymbol{p}) \sim N_K(\boldsymbol{Q}_{\theta}(\boldsymbol{p}), n^{-1} \boldsymbol{\mathcal{K}}_{\theta})
\end{equation}
Here $\theta$ is an unknown parameter to be estimated from the data. The 
covariance matrix $\boldsymbol{\mathcal{K}}_{\theta}$ has entries from the 
covariance function $\kappa_{\theta}(\cdot, \cdot)$ where 
\[
\kappa_{\theta}(p, p') = \frac{p\wedge p' - p p'}{f_{\theta}(Q_{\theta}(p)) f_{\theta}(Q_{\theta}(p'))}, \quad p, p' \in (0,1)
\]

The covariance function $\kappa_{\theta} (\cdot, \cdot)$ initially appears as 
if it will be difficult to evaluate and indeed can be, depending on the 
functional forms of $f_{\theta}$ and $Q_{\theta}$. But where these functions 
belong to a location-scale family, the covariance function can be greatly 
simplified. 
Noting that the covariance is a function of $f_{\theta}(Q_{\theta}(p)) = q_{\theta}(p)$, the QDF, if $\theta = (\mu, \sigma)$ where $\mu$ is a location parameter and $\sigma$ is a scale parameter, then $q_{\theta}(p) = \sigma q(p)$ where $q(p)$ is the QDF of a standard distribution having location 0 and shape 1. \cite{staudte2017shapes} calls this the location invariant and scale equivariant property. For location-scale distribution families, this greatly simplifies the form of $\kappa_{\theta}(\cdot,\cdot)$ so that it becomes
\[
    \kappa_{\theta}(p, p') = \sigma^{2}[p\wedge p' - p p'] q(p)q(p') 
\] 
Thus for estimating $\theta$ one only needs to estimate $\sigma^2$. A special 
case of the QGP for QM of a location-scale family is the normal QGP which we 
define and explore below.

\subsection{Normal QGP}

If the quantiles in a dataset 
$(\hat{\boldsymbol{Q}}_n(\boldsymbol{p}),\boldsymbol{p})$ are assumed to be 
calculated from a normal distribution $N(\mu, \sigma^2)$, and the desire is to 
estimate the parameters $\mu$ and $\sigma$ by modeling the quantiles according 
to the QGP in (\ref{eq:qgp}), then one may use the model (\ref{eq:norm_qgp}).

\begin{equation}
    \label{eq:norm_qgp}
\hat{\boldsymbol{Q}}_n(\boldsymbol{p}) \sim N_K 
\left( \mu + \sigma \boldsymbol{\Phi}^{-1}(\boldsymbol{p}),\frac{\sigma^2}{n} 
\boldsymbol{\Psi} \right)
\end{equation}
Here $\boldsymbol{\Phi}^{-1}(\boldsymbol{p}) = (\Phi^{-1}(p_1), ..., 
\Phi^{-1}(p_K))$ is a known vector since $\boldsymbol{p}$ is given, and 
$\boldsymbol{\Psi}$ is a known $K\times K$ matrix with $ij^{th}$ entry

\[
    \Psi_{ij} = \frac{2 \pi (p\wedge p' - p p')}{\text{exp}\{-\frac{1}{2}[\Phi^{-1}(p)^2 + \Phi^{-1}(p')^2]\}}.
\]
Model (\ref{eq:norm_qgp}) can then be viewed as an atypical normal linear 
regression model. Two aspects that make this model atypical relative to the 
standard linear regression model are that the slope parameter $\sigma$ must be 
greater than 0 and that $\sigma$ is both a regression coefficient and part of 
the variance.
% making $\sigma$ unidentifiable.
Note also that the sample size $n$ is included as part of the variance. If $n$ 
is known, then it may be multiplied through $\Psi$ and forgotten. If unknown, 
it can then be accounted for as an unknown part of the variance.

>>>>

%% The amssymb package provides various useful mathematical symbols
% \usepackage{amssymb}
%% The amsmath package provides various useful equation environments.
% \usepackage{amsmath}
\usepackage{amsmath,amsthm,amssymb,scrextend,mathtools}
\usepackage{hyperref}
\usepackage{newtxmath} %for making indicator function
% \RequirePackage[colorlinks,citecolor=blue,urlcolor=blue,backref=page,backref=page]{hyperref}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

@@

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}
\usepackage{natbib}

@@

% \begin{frontmatter}

@@

\title{Supplementary Material} %% Article title

@@

% \author[label1]{Spencer Wadsworth} %% Author name
% \author[label1]{Jarad Niemi}

%% Author affiliation
% \affiliation[label1]{organization={Iowa State Univerity, Department of Statistics},%Department and Organization
%             addressline={2438 Osborn Dr}, 
%             city={Ames},
%             postcode={50011}, 
%             state={Iowa},
%             country={USA}}

@@

%%Research highlights
% \begin{highlights}
% \item Research highlight 1
% \item Research highlight 2
% \end{highlights}
% 
%% Keywords

% 
% \end{frontmatter}

@@

\Huge
\begin{center}
  Supplementary Material
\end{center}
\normalsize

Here we include a simulation study similar to the first study in section (??).
% \ref{sec:simulation_analyses}. 
However, instead of simulating data from a normal distribution, data is 
simulated 
from an exponential distribution with parameter $\lambda = 4$. The same prior 
distribution for $\lambda$ was used in fitting QGP, ORD, and IND models, and the 
same prior distribution was assigned to $n$ in QGP and ORD models and 
$1/\sigma_{\rho}$. The prior for $\lambda$ was 
$\pi(\lambda) \sim N(0, 7^2)\vmathbb{1}\{\lambda > 0\}$, the prior for $n$ was
$\pi(n) \sim N(0, 3000^2)\vmathbb{1}\{n > 0\}$, and the prior for
$1/\sigma_{\rho}$ was
$\pi(1/\sigma_{\rho}) \sim N(0, 3000^2)\vmathbb{1}\{1/\sigma_{\rho} > 0\}$.

Figure \ref{fig:exp_par_fit} shows examples of posterior sample density plots 
for $\lambda$ and $n$. QGP and ORD show similar distributions which tend to be 
a bit wider than the posterior distributions from the IND model. The left side 
of figure \ref{fig:exp_cov_dists} shows the percent coverage of the 500 
simulation replicates for $\lambda$ and $n$. Here, the IND model coverage is 
much closer to the nominal level than in the normal distribution case of the 
previous section. The right side of figure \ref{fig:exp_cov_dists} shows the 
average UWD1, TV, and KLD distances for the five QM methods. Again the 
parametric models tend to perform better than the non-parametric SPL and KDE 
methods with the SPL and KDE performing better as $n$ and $K$ increase.

\begin{figure}

\centering
% \begin{subfigure}
  \centering
  \hspace{-.8cm}
  \includegraphics[width=.94\linewidth]{Images/exponential_par_fit.png}
  % \caption{A subfigure}
  % \label{fig:sub1}
% \end{subfigure}%
% \begin{subfigure}
  \centering
  \hspace{-.9cm}
  \includegraphics[width=1.05\linewidth]{Images/exponential_n_fit.png}
  % \caption{A subfigure}
  % \label{fig:sub2}
% \end{subfigure}
\caption{Density plots of posterior distribution samples for the exponential 
parameters by QM for QGP, ORD, and IND models. QM was done on estimated 
quantiles from a exponential distribution with parameter $\lambda = 4$. The 
posterior densities are for $\lambda$ (left) and sample size $n$ (right). Plots 
are faceted by true sample size $n \in \{50, 150, 500, 1{,}000\}$ ($y$-axis) 
and number of quantiles $K \in \{7, 15, 23\}$ ($x$-axis). Vertical lines 
(black) show the value of the true parameter.}
\label{fig:exp_par_fit}
\end{figure}

\begin{figure}[hbt!]
% \begin{subfigure}
  \includegraphics[width=\linewidth]{Images/exponential_coverage.png}
  % \caption{}
  % \label{MLEDdet}
% \end{subfigure}\hfill % <-- "\hfill"
% \begin{subfigure}
  \includegraphics[width=\linewidth]{Images/exponential_dists.png}
  % \caption{}
  % \label{energydetPSK}
% \end{subfigure}
\caption{Posterior coverage (left) calculated as the percentage of times the 
true parameter fell within the modeled 90\% credible interval over the 500 
replications. Coverage is faceted by the exponential parameter 
$\lambda$ and $n$ with $K \in \{3, 7, 15, 23, 50\}$, and by increasing sample 
size ($x$-axis). The five models QGP, ORD, QGPN, ORDN, and IND are colored as 
shown the legend. The horizontal line (black) is at the nominal 90\% level. 
Only QGP and ORD appear for the parameter $n$ as they are the only two which 
estimate an unknown $n$.
Distance between the true distribution and the estimated QM predictive 
distribution (right) averaged over the 500 replications. Distances include 
UWD1, TV, and KLD for $K \in \{3, 7, 15, 23, 50\}$, and by increasing sample 
size ($x$-axis).}
\label{fig:exp_cov_dists}
\end{figure}

\begin{figure}[hbt!]
\centering
%\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.49\linewidth]{Images/quants_lp.png}
  % \caption{A subfigure}
  % \label{fig:sub1}
%\end{subfigure}%
%\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.49\linewidth]{Images/dens_lp.png}
  % \caption{A subfigure}
  % \label{fig:sub2}
%\end{subfigure}
\caption{QM fits of $K=23$ quantiles by KDE, SPL, IND, and QGP for $n \in \{50, 150, 500, 1{,}000\}$. The quantiles were sampled from the Laplace distribution $La(0,1)$. The quantile fits (left) show the true quantiles (black) with either the QM fit line (grey) or the credible intervals of 50\% (red) and 95\% (pink). 
The estimated PDF plots (right) show the true PDF (black) with either a the QM estimated PDF (grey) or the credible intervals of 50\% (red) and 95\% (pink).}
\label{fig:lp_fits}
\end{figure}

\begin{figure}[hbt!]
\centering
%\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.49\linewidth]{Images/quants_gmix.png}
  % \caption{A subfigure}
  % \label{fig:sub1}
%\end{subfigure}%
%\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.49\linewidth]{Images/dens_gmix.png}
  % \caption{A subfigure}
  % \label{fig:sub2}
%\end{subfigure}
\caption{QM fits of $K=23$ quantiles by KDE, SPL, IND, and QGP for $n \in \{50, 150, 500, 1{,}000\}$. The quantiles were sampled from the two component normal mixture distribution $w N(-1, 0.9) + (1-w)N(1.2, .6)$ where $w = 0.35$. The quantile fits (left) show the true quantiles (black) with either the QM fit line (grey) or the credible intervals of 50\% (red) and 95\% (pink). 
The estimated PDF plots (right) show the true PDF (black) with either a the QM estimated PDF (grey) or the credible intervals of 50\% (red) and 95\% (pink). }
\label{fig:gmix_fits}
\end{figure}
